Generating Datasets about Building Designs \\Based on a Multilayer Building Ontology

### 2015-05-01 article-01-introduction.tex
Version 1.5

\section{Introduction}
\label{sec:Introduction}

@@

CityGML \cite{kolbe2005citygml}, a surface model for buildings, can be useful to visitors for 
indoor navigation. IFC contains complementary information about building design, structure, 
and equipment for a wider range of stakeholders and use cases. Moreover, IFC data can be one source 
when generating CityGML models. % GityGML has also been converted into OWL.

\TODO{Update the description of sections}

In the rest of this paper, we describe the conversions to \ifcowl{} and \ifcrdf{}, present the deployment of ontologies, example datasets and the converter, and finish with summary and future work.

>>>>

\section{Introduction}
\label{sec:introduction}

@@

% CityGML \cite{kolbe2005citygml}, a surface model for buildings, can be useful to visitors for 
% indoor navigation. IFC contains complementary information about building design, structure, 
% and equipment for a wider range of stakeholders and use cases. Moreover, IFC data can be one source 
% when generating CityGML models. % GityGML has also been converted into OWL.

% \TODO{Update the description of sections}

% In the rest of this paper, we describe the conversions to \ifcowl{} and \ifcrdf{}, present the deployment of ontologies, example datasets and the converter, and finish with summary and future work.

In the Sect. \ref{sec:ifcOWL}, we describe the principles of IFC-EXPRESS-to-OWL conversion. Sect. \ref{sec:ifcRDF} provides some sample fragments of IFC-SPF-to-RDF conversion; where SPF (STEP Physical File) defined in ISO 10303-21 \cite{ISO10303-21} is the main exchange format of IFC data. Sect. \ref{sec:deployment} includes short descriptions about public access to the \textbf{IFC2LD} converter and the resulting ifcOWL ontologies.

### 2015-05-01 article-02-main.tex
Version 1.5

\subsubsection{Entity Datatypes} correspond to classes in object-oriented languages. Each type may have one super\-type. Thus, an entity data\-type should be converted into a subclass of its super\-class or of \name{expr:\-Entity\-Class}. For instance, a type with attributes \name{Profiles : SET [2:?] OF IfcProfileDef} and \name{Label : OPTIONAL IfcLabel} is defined as follow:

%\begin{lstlisting}[caption={Entity data\-types}, label=lst:ifcOWL-defined-types]
\begin{lstlisting}
ifc:IfcCompositeProfileDef a owl:Class ; rdfs:subClassOf ifc:IfcProfileDef ;
  # only for Standard, Extended layers
    rdfs:subClassOf [ a owl:Restriction; owl:onProperty ifc:profiles;
                      owl:allValuesFrom ifc:SET_2_UNLIMITED_IfcProfileDef ] ;
    rdfs:subClassOf [ a owl:Restriction; owl:onProperty ifc:label;
                      owl:allValuesFrom ifc:Label ] ;
  # only for Extended layer
    rdfs:subClassOf [ a owl:Restriction; owl:onProperty ifc:profiles;
                      owl:cardinality "1"^^xsd:integer ] ;
    rdfs:subClassOf [ a owl:Restriction; owl:onProperty ifc:label;
                      owl:minCardinality "0"^^xsd:integer ] ;
    rdfs:subClassOf [ a owl:Restriction; owl:onProperty ifc:label;
                      owl:maxCardinality "1"^^xsd:integer ] .
\end{lstlisting}

The minimum property cardinality of \name{ifc:label} is 0 because it is optional. The maximum cardinality of any property is 1. If an attribute type is a set (or a list), then a new specific aggregation must be defined, for example, \name{SET\_\-2\_\-UNLIMITED\_\-IfcProfileDef} is created to specify property \name{ifc:profiles}. This approach allows to represent lists, sets and arrays in the same manner and provides possibilities to constraint the attribute cardinalities.

@@

\subsubsection{Header Section.} IFC data is exchanged as \emph{STEP physical files} (SPF) format of which is defined in ISO 10303-21 \cite{ISO10303-21}. A SPF header section includes a list of STEP-specific metadata entities with types \name{FileDescription}, \name{FileName} and \name{FileSchema}. %described in \cite{buildingSMART:ifc-header}. 
They are exported as objects of STEP entities, some attributes are duplicated as Dublin Core (DC) metadata annotations, for example:

@@

\subsubsection{Data Section.} In a SPF data section each entity is identified with a line number.
%, an entity data\-type and a list of values. 
All entities of \name{Ifc\-Root}-derived types have a GUID in IFC. They are converted into URI resources with format \name{GUID\_\emph{\textless{}guid\textgreater{}}} where \emph{guid} is in the hexadecimal string representation (not including the character \name{\$}). Regarding to anonymous IFC entities, it can be chosen whether to export them as blank nodes or create them a URI in a format \name{LINE\_\emph{\textless{}lineNumber\textgreater{}}} (in a user defined namespace). Note that line numbers are used only for linking entities inside a SPF, and they are unstable. Naming is required when the dataset is used together with OWL 2 EL/QL as these profiles do not allow anonymous individuals. Below is an example of IFC data exported to an RDF dataset compatible with OWL 2 RL:

@@

\section{Deployment}

The \textbf{IFC2LD converter}, \ifcowl{} ontologies, and example \ifcrdf{} datasets (together with the original IFC files) are available for use from the addresses:

% \vspace{0.5\topsep}
\noindent
\begin{scriptsize}
\begin{center}
    \def\arraystretch{1.2}
    \setlength{\tabcolsep}{6pt}
    \begin{tabularx}{0.90\textwidth}{|l|X|}
        \hline
        Ontologies: & \texttt{http://drumbeat.cs.hut.fi/owl/} \\
        \hline
        Example files & \\
        - IFC data: & \texttt{http://drumbeat.cs.hut.fi/ifc/} \\
        - RDF datasets: & \texttt{http://drumbeat.cs.hut.fi/rdf/} \\
        \hline
        IFC2LD Converter & \\
        - online service: & \texttt{http://drumbeat.cs.hut.fi/home/} \\
        - download binary: & \texttt{http://drumbeat.cs.hut.fi/ifc2ld/} \\
        - source code: & \texttt{https://github.com/Web-of-Building-Data/Ifc2Rdf} \\
        \hline
    \end{tabularx}
\end{center}
\end{scriptsize}

@@

We have presented \ifcowl{}, a multilayer ontology of building designs based on the widely adopted IFC standard. The ontology is 
derived from the IFC schema with the \ifcld{} converter that can also produce RDF datasets from IFC files.

The \ifcowl{} ontologies and \ifcrdf{} datasets are currently used in a variety of Linked Data oriented use cases \cite{torma2013semantic} developed in the DRUMBEAT project. We are exploring the possibilities to use the \ifcld{} converter in the generation of metadata for \ifcrdf{} datasets and, in some cases, also linksets between datasets.

>>>>

\subsubsection{Entity Datatypes} correspond to classes in object-oriented languages. Each type may have one super\-type. Thus, an entity data\-type should be converted into a subclass of its super\-class or of \name{expr:\-Entity\-Class}.

The minimum cardinality of an optional entity property is \name{0}. The maximum cardinality is always \name{1}. If an attribute type is set, list, or array, then a new aggregation data\-type is defined, for example, \name{ifc:\-SET\_\-2\_\-UNLIMITED\_\-IfcProfileDef} was created for property \name{Profiles : SET [2:?] OF IfcProfileDef}. This approach allows to represent all multiple attributes in the same manner and provides possibilities to constrain the cardinalities, even when the minimum cardinality is not \name{0} or \name{1}.

% Below is an example of definition of a type with attributes \name{Profiles : SET [2:?] OF IfcProfileDef} and \name{Label : OPTIONAL IfcLabel}:

%\begin{lstlisting}[caption={Entity data\-types}, label=lst:ifcOWL-defined-types]
\begin{lstlisting}
ifc:IfcCompositeProfileDef a owl:Class ; rdfs:subClassOf ifc:IfcProfileDef ;
  # only for the Standard, Extended layers
    rdfs:subClassOf [ a owl:Restriction; owl:onProperty ifc:profiles;
                      owl:allValuesFrom ifc:SET_2_UNLIMITED_IfcProfileDef ] ;
    rdfs:subClassOf [ a owl:Restriction; owl:onProperty ifc:label;
                      owl:allValuesFrom ifc:Label ] ;
  # only for the Extended layer
    rdfs:subClassOf [ a owl:Restriction; owl:onProperty ifc:profiles;
                      owl:cardinality "1"^^xsd:integer ] ;
    rdfs:subClassOf [ a owl:Restriction; owl:onProperty ifc:label;
                      owl:minCardinality "0"^^xsd:integer ] ;
    rdfs:subClassOf [ a owl:Restriction; owl:onProperty ifc:label;
                      owl:maxCardinality "1"^^xsd:integer ] .
\end{lstlisting}

@@

As mentioned in Sect. \ref{sec:introduction}, IFC data is exchanged in STEP Physical Files (SPF). The IFC-SPF-to-RDF conversion is aimed to produce \ifcrdf{} datasets including \emph{all} information stored in SPFs and compatible with any \ifcowl{} layer.

\subsubsection{Header Section} includes STEP-specific metadata as normal data-section entities. Their types -- \name{FileDescription}, \name{FileName}, and \name{FileSchema} -- are defined in STEP. They are exported as normal entities, but some attributes of them are duplicated as Dublin Core (DC) metadata annotations, for example:

@@

\subsubsection{Data Section} includes entities, each of which is identified with a line number.
%, an entity data\-type and a list of values. 
All entities of \name{Ifc\-Root}-derived types have a GUID in IFC. They are converted into URI resources with format \name{GUID\_\emph{\textless{}guid\textgreater{}}} where \emph{guid} is in the hexadecimal string representation (not including the character \name{\$}). Regarding to anonymous IFC entities, it can be chosen whether to export them as blank nodes or create them a URI in a format \name{LINE\_\emph{\textless{}lineNumber\textgreater{}}} (in a user defined namespace). Note that line numbers are used only for linking entities inside a SPF, and they are unstable. Naming is required when the dataset is used together with OWL 2 EL/QL as these profiles do not allow anonymous individuals. Below is an example of IFC data exported to an RDF dataset compatible with OWL 2 RL:

@@

\section{Implementation and Deployment}
\label{sec:deployment}

The algorithm of generating \ifcowl{} ontologies and \ifcrdf{} datasets, including the principles described in Sect. \ref{sec:ifcOWL}-\ref{sec:ifcRDF}, are implemented in our Java program, namely \textbf{IFC2LD converter}. One can use its command-line interface, web interface or application programming interface for converting any IFC schema or data file into \ifcowl{} or \ifcrdf{}. By changing the configuration file, one can also force the converter to adapt various OWL profiles starting from OWL Lite.

The \textbf{IFC2LD converter}, \ifcowl{} ontologies including different layers for different IFC schema versions, and some sample \ifcrdf{} datasets (together with the original IFC files) are available on the Internet (Table \ref{tab:ifc2ld-web}).

\begin{table}
    \caption{Contents of ifcOWL layers}
    \label{tab:ifc2ld-web}
    \centering
    \scriptsize
    \def\arraystretch{1.2}          % vertical space
    \setlength{\tabcolsep}{6pt}     % horizontal space
    \begin{tabularx}{0.90\textwidth}{|l|X|}
        \hline
        Ontologies: & \texttt{http://drumbeat.cs.hut.fi/owl/} \\
        \hline
        Example files & \\
        - IFC data: & \texttt{http://drumbeat.cs.hut.fi/ifc/} \\
        - RDF datasets: & \texttt{http://drumbeat.cs.hut.fi/rdf/} \\
        \hline
        IFC2LD Converter & \\
        - online service: & \texttt{http://drumbeat.cs.hut.fi/home/} \\
        - download binary: & \texttt{http://drumbeat.cs.hut.fi/ifc2ld/} \\
        % - source code: & \texttt{https://github.com/Web-of-Building-Data/Ifc2Rdf} \\
        \hline
    \end{tabularx}
\end{table}

@@

In this paper, we proposed a new approach and principles of creating \ifcowl{}, a multilayer ontology of building designs based on the widely adopted IFC standard. The ontology is 
derived from the IFC schema with the \ifcld{} converter that can also produce RDF datasets from IFC files.

The \ifcowl{} ontologies and \ifcrdf{} datasets are currently used in a variety of Linked-Data-oriented use cases \cite{torma2013semantic} which are developed in the DRUMBEAT project. We are exploring the possibilities to use the \ifcld{} converter in the terms of generation of metadata for linksets between \ifcrdf{} datasets.
